---
title: "Thailand Worst Plastic Polluters"
description: |
  Companies contributing to the greatest portions of plastic polution in Thailand.
author:
  - name: Joe Walderman
    url: {}
date: 02-21-2021
output:
  distill::distill_article:
    self_contained: false
    code_folding: hide
    theme: lumen
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(tidytuesdayR)
library(glue)
library(treemapify)
library(ggtext)
library(extrafont)
```

 #breakfreefromplastic is a global movement with support from over 12,000 organizations and individuals around the world. The goal of the movement is to reduce single-use plastics and work towrds solutions for the plastic pollution crisis.
 
 <br>
 
 In 2020, a collection of 905 volunteers collected 38,580 pieces of plastic from 115 companies in Thailand.

```{r}
plastics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-01-26/plastics.csv')




##use data for pet,pp and ps type plastics only

# Replace null with unbranded
plastics_clean <- plastics %>%
  # Replace null with Unbranded in Parent Company
  mutate(parent_company = ifelse(tolower(parent_company) == 'null', 
                                 'Unbranded', 
                                 parent_company)) %>%
  # Remove Grand Total in Parent Company
  filter(!(tolower(parent_company) == "grand total")) %>%
  # Standardize Nestlé vs Nestle
  mutate(parent_company = ifelse(parent_company == 'Nestle',
                                 'Nestlé',
                                 parent_company)) %>%
  # Make long by plastic type
  pivot_longer(empty:grand_total, 
               names_to = 'plastic_type', 
               values_to = 'count',
               values_drop_na = TRUE) %>%
  # Remove 0 counts
  filter(count != 0)


plastics_philippines <- 
  plastics %>% 
  filter(year == 2020, !parent_company %in% c("Unbranded", "NULL", "null"), country == "Thailand") %>%    
  mutate(parent_company = case_when(grand_total < 25 ~ "112 Other Companies",
                                    TRUE ~ parent_company)) %>% 
    add_count(parent_company)

#Total plastic objects in The Philippines
plastics_philippines %>% 
  summarise(total = sum(grand_total)) -> philippines_total

#Volunteers in The Philippines
philippines_volunteers <- plastics_philippines[1, 14]

#Companies in The Philippines
philippines_companies <- nrow(plastics_philippines)

#Plot colors  
bigger <- "midnightblue"
middle <- "slateblue3"
small <- "skyblue4"
other <- "magenta4"
text <- "black"

#Plot  
plastics_philippines %>%  
    group_by(parent_company) %>%  
    summarise(grand_total = sum(grand_total)) %>% 
    mutate(parent_company = paste0(parent_company, "\n(", grand_total, ")"),
           group = case_when(grand_total > 250 ~ glue("{bigger}"),
                             grand_total > 100 ~ glue("{middle}"),
                             grand_total == 243 ~ glue("{other}"),
                             TRUE ~ glue("{small}"))) %>% 
    ggplot(aes(area = grand_total, label = parent_company, fill = group)) +
    geom_treemap() +
    geom_treemap_text(color = "white", reflow = TRUE) +
    scale_fill_identity() +
    labs(title = "**Worst Plastic Polluting Companies in Thailand**<br>",
         caption = "2 Feb 2021 #TidyTuesday dataset || Joe Walderman || Inspired by @luisfreii") +
    theme(plot.margin = unit(c(2, 2, 2, 2), "lines"),
          plot.title = element_markdown(hjust = .5, size = 14, color = text),
          plot.subtitle = element_markdown(hjust = .5, color = text),
          plot.caption = element_markdown(hjust = .5, color = text)
          )

